---
title: "Evolution of Metadata-driven Analysis Results"
format: html
toc: true
---

#
#load packages
```{r}
library(dplyr)
library(pharmaverseadam)
library(cards)
```

```{r}
data(adae)
data(adsl)
```


# 1: metadata and base code

Starting with the most simplistic approach. Here we want to demonstrate how we can use a bespoke metadata structure to drive commonly used R packages like `dplyr`. 
```{r}
metadata <- list(
    result_id = "AE_SUMMARY_SEVERE",              # Unique identifier for the result
    analysis_purpose = "Safety",                  # Purpose of the analysis
    analysis_dataset = "ADAE",                    # Source dataset
    analysis_population = "Safety Population",
    parameter = "Subjects with Severe Adverse Events", # Parameter being summarised
    analysis_variable = c("AEBODSYS", "AEDECOD"),     # Variables used in the analysis
    analysis_value = c("COUNT", "PERCENTAGE"),        # Values displayed
    treatment_variable = "TRT01A",                # Grouping variable
    filter_condition = "AESEV == 'SEVERE'",        # Applied filter
    display_title = "Summary of Severe Adverse Events by Treatment Arm",
    display_type = "Table",                       # ARS Display attribute
    notes = "Includes treatment-emergent severe adverse events only"
)
```

```{r}
library(rlang)

# Pull names from metadata so we don't hardcode them below
trt_var   <- metadata$treatment_variable
anal_vars <- metadata$analysis_variable

# 1) Filter ADAE per ARS metadata (e.g., severe AEs)
adae_f <- adae %>%
    filter(!!parse_expr(metadata$filter_condition))

# 2) Denominators N by treatment arm from ADSL (Safety population)
denom <- ADSL %>%
    filter(SAFFL == "Y") %>%                      # adjust if your pop flag differs
    distinct(USUBJID, .keep_all = TRUE) %>%       # one row per subject
    count(!!sym(trt_var), name = "N") %>%         # N per treatment arm
    rename(treatment_group = !!sym(trt_var))

# 3) Subject counts per SOC/PT per treatment (COUNT) and PERCENTAGE = COUNT/N*100
ars_results <- adae_f %>%
    # subject-level dedup within grouping (ensures "subjects with ≥1 AE", not event counts)
    distinct(USUBJID, !!sym(trt_var), across(all_of(anal_vars))) %>%
    count(!!sym(trt_var), across(all_of(anal_vars)), name = "COUNT") %>%
    rename(treatment_group = !!sym(trt_var)) %>%
    left_join(denom, by = "treatment_group") %>%
    mutate(PERCENTAGE = if_else(N > 0, 100 * COUNT / N, NA_real_)) %>%
    # 4) Attach a few ARS fields as columns (same values on all rows)
    mutate(
        result_id           = metadata$result_id,
        analysis_purpose    = metadata$analysis_purpose,
        analysis_dataset    = metadata$analysis_dataset,
        analysis_population = metadata$analysis_population,
        parameter           = metadata$parameter,
        analysis_value      = paste(metadata$analysis_value, collapse = ", "),
        treatment_variable  = trt_var,
        filter_condition    = metadata$filter_condition,
        display_title       = metadata$display_title,
        display_type        = metadata$display_type
    ) %>%
    relocate(
        result_id, analysis_purpose, analysis_dataset, analysis_population,
        parameter, treatment_variable, treatment_group,
        all_of(anal_vars), COUNT, N, PERCENTAGE,
        analysis_value, filter_condition,
        display_title, display_type
    )

```



# 2: metadata and packaged code
- simple metadata, maybe add subgroups
- use a package like cards

# 3: Scaling in your org: standardised metadata and packaged code
- CDISC ARS concepts
- demo internal tool?